---
title: "Accuracy and call rate vs. q"
author: "Corresponding author"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggnewscale) # Enables multiple colour- and fill-scales
library(latex2exp)
library(patchwork)
source("00_functions_and_global_definitions.R")

# The palette with grey:
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

plot_title_size <- 7
axis_title_size <- 7
axis_ticks_size <- 7
legend_title_size <- 7
legend_text_size <- 7
```

```{r Reading data}
dd_GT <- readRDS(file.path("..", "data", "00_dd_GT.rds"))
dd_GT_2nd <- readRDS(file.path("..", "data", "00_dd_GT_2nd.rds"))
true_profiles_2nd <- readRDS(file.path("..", "data", "00_true_profiles_2nd.rds"))
# We will treat the different runs in dd_GT as independent, i.e. simply consider
# a new run as new individuals which we encode by interaction(Run, ID, drop=T)
dd_GT$ID <- interaction(dd_GT$ID, dd_GT$Run, drop = TRUE)

dd_GT_both <- dd_GT_2nd |> bind_rows(dd_GT[names(dd_GT_2nd)])
```

## Scheme with the train data's list entrances (low range of DNA amounts)

Letter |      Train     | Test (low)
------ | -------------- | ----------
A      | 250, 125, 62.5 | 50, 31.25
B      | 50, 31.25      | 50, 31.25
C      | 125, 25        | 50, 31.25
D      | 50, 25         | 50, 31.25
E      | 31.25, 25      | 50, 31.25
F      | 50             | 50, 31.25
G      | 31.25          | 50, 31.25
H      | 25             | 50, 31.25
I      | 1000           | 50, 31.25
J      | 1000, 25       | 50, 31.25

```{r}
# NOTE: the name "...9individuals_low.rds" is a bit off now, since we don't sample nine
# individuals for fitting, but 75% corresponding to 9 individual for 31.25pg,
# but not in any other dilutions, where it will be more.

# NOTE: the difference between the objects named
# 01_obj_crossval_NI_1K_9individuals_low.rds
# and
# 01_obj_crossval_NI_1K_9individuals_low_all.rds
# is that the latter contains ALL probability estimates for the test data, not just
# the probability estimates for the points with estimated values lower than the highest WC.
# This also makes the object much larger in size.
# They are both created in 01_crossval_custom.R by commenting in and out the relevant lines.
# dd_val_dils <- readRDS(file.path("..", "data", "01_obj_crossval_sqrt_NI_1K_9individuals_low_all.rds"))
dd_val_dils <- readRDS(file.path("..", "data", "01_obj_crossval_sqrt_1K_9individuals_low_high_all.rds"))
dd_val_dils <- dd_val_dils[1]
names(dd_val_dils) <- "D"

floor_dec <- function(x, level=1) round(x - 5*10^(-level-1), level)
ceiling_dec <- function(x, level=1) round(x + 5*10^(-level-1), level)

q_u <- 0.75
q_l <- 0.25
names_col <- c(paste0("Q", q_l*100), "Median", paste0("Q", q_u*100))
```


# LOW
## Accuracy and call rate vs. q for the cv winner fit testing on low dilutions
```{r}
df_acc <- lapply(dd_val_dils$D, function(x) {
  # Including all points from each of the 1000 cross-validation-iterations will
  # give too huge a data frame, which isn't suited for fitting.
  # We certainly want the q-values of the WCs and then the q-value immediately
  # after the largest q for the WCs.
  q <- c(x$p_WC, min(x$p_NN[x$p_NN > max(x$p_WC)]))
  WC <- sapply(q, function(y) {sum(x$p_WC >= y)})
  p_all <- c(x$p_WC, x$p_NN)
  NC <- sapply(q, function(y) {sum(p_all < y)})
  
  q <- floor_dec(q, 4)
  data.frame("q" = q,
             "Accuracy" = ((x$N_test - WC - NC) / (x$N_test - NC))*100)
}) |> bind_rows(.id = "Line") ##|> 
  # mutate(Line = as.factor((80+(as.numeric(Line)-1)*(100-80)/(1000-1))/100))

# In order to compute a sensible accuracy for each q-value, we make a sliding
# window and extract the points with estimated probabilities in the vicinity of
# each q-value, and then compute the max, min, and median for these points.
qq <- seq(0.5, 1, 0.0005)
acc <- matrix(NA, nrow = length(qq), ncol = 3)
acc[1,] <- quantile(filter(df_acc, q < 0.505)$Accuracy, probs = c(q_l, 0.5, q_u))
acc[length(qq),] <- quantile(filter(df_acc, q >= 0.995)$Accuracy, probs = c(q_l, 0.5, q_u))
for (i in 2:(length(qq)-1)) {
  acc[i,] <- quantile(filter(df_acc, q >= qq[i]-0.005, q < qq[i]+0.005)$Accuracy, probs = c(q_l, 0.5, q_u))
}
colnames(acc) <- names_col
acc <- as.data.frame(acc)
acc$q <- qq
acc <- acc |> pivot_longer(cols = !q, names_to = "Line", values_to = "Accuracy") |> 
  mutate(Method = "SMLR") |> 
  select(Method, q, Line, Accuracy)
# NOTE: the accuracy is NOT a strictly increasing function in q, since we defined it as
# Accuracy = (Total calls - NCs - WCs) / (Total calls - NCs)
# so it actually only increases when the WCs decrease.
# When q increases, the NCs will increase and the WCs will decrease, but the NCs
# may very well take steps up without the WCs taking steps down, in which case
# the Accuracy can decrease a bit.
# Ideally, the WCs goes to zero as the first points turn into NCs, and in that
# case, the Accuracy is strictly increasing, but NOT generally.

df_acc_m <- lapply(dd_val_dils$D, function(x) {
  data.frame("Accuracy_HSG" = ((x$N_test - x$HSG_WC - x$HSG_NN) / (x$N_test - x$HSG_NN))*100)
}) |> bind_rows()
df_acc_m <- data.frame("Accuracy" = c(min(df_acc_m$Accuracy_HSG),
                                      quantile(df_acc_m$Accuracy_HSG, probs = c(q_l, 0.5, q_u)),
                                      max(df_acc_m$Accuracy_HSG)),
                       "Method" = rep("HSG", 5),
                       "Line" = c("Min", names_col, "Max"),
                       "x1" = 0.5,
                       "x2" = 1) |> 
  pivot_wider(values_from = "Accuracy", names_from = "Line")
df_acc_stats <- df_acc_m |> select(Method, Q25, Median, Q75, x1, x2) |> 
  # pivot_longer(cols = c(x1, x2), values_to = "q", names_to = NULL) |> 
  pivot_longer(cols = c(Q25, Median, Q75), values_to = "Accuracy", names_to = "Line")


(p_acc <- ggplot() + 
  # geom_line(aes(x=q, y=Accuracy, alpha=Line), colour="grey", show.legend = F, data = df_acc) +
  geom_line(aes(x=q, y=Accuracy, colour=Line), show.legend = F, data = df_acc) +
  scale_colour_manual(values = rep("grey",1000)) +
  new_scale_color() +
  geom_rect(aes(xmin=x1, xmax=x2, ymin=Min, ymax=Max, fill=Method), data=df_acc_m, alpha=0.3, show.legend = F) +
  scale_fill_manual(breaks="HSG",
                    values= "#F8766D",
                    labels="HSG") +
  # geom_rect(aes(xmin=x1, xmax=x2, ymin=Q25, ymax=Q75, fill=Method), data=df_acc_m, alpha=0.6) +
  geom_smooth(aes(x=q, y=Accuracy, lty=Line, colour=Method), se=F, linewidth=0.5, data = acc) +
  geom_segment(aes(x=x1, y=Accuracy, xend=x2, yend=Accuracy, colour=Method, lty=Line), data=df_acc_stats) +
  scale_colour_manual(values = c("#F8766D", "black"),
                      breaks = c("HSG", "SMLR")) +
  scale_linetype_manual(values = c("dotted", "dashed", "dotted"),
                        breaks = c("Q75", "Median", "Q25")) +
  coord_cartesian(xlim = c(0.5, 1), ylim = c(98.75, 100), expand = F) +
  scale_x_continuous(breaks = seq(0.5, 1, 0.1),
                     labels = c(seq(0.5, 0.9, 0.1), 1)) +
  theme_bw() +
  ggtitle("Fit: 25pg, 50pg\nTest: 31.25pg, 50pg") +
  theme(plot.title = element_text(size = plot_title_size),
        axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_ticks_size),
        legend.position = "none"))

# ggsave(filename = file.path("..", "article-FSIGEN", "accuracy_vs_q_low.pdf"), plot = p_acc,
#        device = "pdf", width = 9, height = 7, units = "cm", dpi = 1000)
```


```{r}
y_min <- 85 # minimal call rate to be displayed

df_cr <- lapply(dd_val_dils$D, function(x) {
  q <- c(x$p_WC, x$p_NN)
  q <- floor_dec(q[order(q)], 4)
  data.frame("q" = q,
             "Call_rate" = ((x$N_test - 0:(length(q) - 1)) / x$N_test)*100)
}) |> bind_rows(.id = "Line")


qq <- seq(0.5, 0.95, 0.005)
qq2 <- seq(0.96, 1, 0.001)
cr <- matrix(NA, nrow = length(c(qq, qq2)), ncol = 3)
cr[1,] <- quantile(filter(df_cr, q < 0.505)$Call_rate, probs = c(q_l, 0.5, q_u))
cr[nrow(cr),] <- quantile(filter(df_cr, q >= 0.9995)$Call_rate, probs = c(q_l, 0.5, q_u))
for (i in 2:(length(qq))) {
  cr[i,] <- quantile(filter(df_cr, q >= qq[i]-0.005, q < qq[i]+0.005)$Call_rate, probs = c(q_l, 0.5, q_u))
}
for (i in 1:(length(qq2)-1)) {
  cr[length(qq)+i,] <- quantile(filter(df_cr, q >= qq2[i]-0.0005, q < qq2[i]+0.0005)$Call_rate, probs = c(q_l, 0.5, q_u))
}
colnames(cr) <- names_col
cr <- as.data.frame(cr)
cr$q <- c(qq, qq2)
cr <- cr |> pivot_longer(cols = !q, names_to = "Line", values_to = "Call_rate") |> 
  mutate(Method = "SMLR") |> 
  select(Method, q, Line, Call_rate)
# In reality, the call rate should be strictly decreasing with increasing q, but
# the sliding window makes it possible to have tiny steps up, which is meaningless
# and disturbing. It's not as such visible in the plots, but to smooth the results,
# we reorder and re-map the call rate.
cr <- cr |> group_by(Line) |> arrange(desc(Call_rate), .by_group = T) |> mutate(q = c(qq, qq2))


df_cr_m <- lapply(dd_val_dils$D, function(x) {
  data.frame("Call_rate_HSG" = ((x$N_test - x$HSG_NN) / x$N_test)*100)
}) |> bind_rows()
df_cr_m <- data.frame("Call_rate" = c(min(df_cr_m$Call_rate_HSG),
                                      quantile(df_cr_m$Call_rate_HSG, probs = c(q_l, 0.5, q_u)),
                                      max(df_cr_m$Call_rate_HSG)),
                      "Method" = rep("HSG", 5),
                      "Line" = c("Min", names_col, "Max"),
                       "x1" = 0.65,
                       "x2" = 1) |> 
  pivot_wider(values_from = "Call_rate", names_from = "Line")
df_cr_stats <- df_cr_m |> select(Method, Q25, Median, Q75, x1, x2) |> 
  # pivot_longer(cols = c(x1, x2), values_to = "q", names_to = NULL) |> 
  pivot_longer(cols = c(Q25, Median, Q75), values_to = "Call_rate", names_to = "Line")



(p_cr <- ggplot() + 
  # geom_line(aes(x=q, y=Call_rate, alpha=Line), colour="grey", show.legend = F, data = df_cr) +
  geom_line(aes(x=q, y=Call_rate, colour=Line), show.legend = F, data = df_cr) +
  scale_colour_manual(values = rep("grey",1000)) +
  new_scale_color() +
  geom_rect(aes(xmin=x1, xmax=x2, ymin=Min, ymax=Max, fill=Method), data=df_cr_m, alpha=0.3, show.legend = F) +
  scale_fill_manual(breaks="HSG",
                    values= "#F8766D",
                    labels="HSG") +
  # geom_rect(aes(xmin=x1, xmax=x2, ymin=Q25, ymax=Q75, fill=Method), data=df_acc_m, alpha=0.6) +
  geom_line(aes(x=q, y=Call_rate, lty=Line, colour=Method), data = cr) +
  geom_segment(aes(x=x1, y=Call_rate, xend=x2, yend=Call_rate, colour=Method, lty=Line), data=df_cr_stats) +
  scale_colour_manual(values = c("#F8766D", "black"),
                      breaks = c("HSG", "SMLR")) +
  scale_linetype_manual(values = c("dotted", "dashed", "dotted"),
                        breaks = c("Q75", "Median", "Q25")) +
  guides(lty = guide_legend(position = "inside"),
         colour = guide_legend(position = "inside")) +
  coord_cartesian(xlim = c(0.5, 1), ylim = c(y_min, 100), expand = F) +
  scale_x_continuous(breaks = seq(0.5, 1, 0.1),
                     labels = c(seq(0.5, 0.9, 0.1), 1)) +
  theme_bw() +
  # ggtitle("Fit: 50pg, 25pg\nTest: 50pg, 31.25pg") +
  theme(plot.title = element_text(size = plot_title_size),
        axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_ticks_size),
        legend.title = element_text(size = legend_title_size),
        legend.text = element_text(size = legend_text_size),
        legend.position.inside = c(0.15,0.41),
        legend.spacing.y = unit(1, 'mm'),
        legend.key.spacing = unit(1, 'mm'),
        legend.key.height = unit(4, 'mm'),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")) +
  ylab("Call rate"))

# ggsave(filename = file.path("..", "article-FSIGEN", "call-rate_vs_q_low.bmp"), plot = p_cr,
#        device = "bmp", width = 9, height = 7, units = "cm", dpi = 1000)
```

```{r}
p <- p_acc + p_cr & theme(plot.margin = margin(t=0, r=5.5, b=0, l=5.5, unit = 'pt'))
p[[1]]$theme$plot.margin[4] <- 0*p[[1]]$theme$plot.margin[4]
p[[2]]$theme$plot.margin[2] <- 0*p[[2]]$theme$plot.margin[2]
(p <- p + plot_annotation(theme = theme(plot.margin = margin(t=0, r=1, b=0, l=0, unit = 'pt'))))

ggsave(filename = file.path("..", "article-FSIGEN", "FIG_4-two-column.bmp"),
       plot = p, device = "bmp", width = 7480, height = 2440, units = "px", dpi = 1000)
```

### Summary statistics for low
```{r}
# HSG median accuracy and call rate
(a_med_hsg <- filter(df_acc_m, Method == "HSG")$Median)
(filter(df_acc_m, Method == "HSG")$Max)
(c_med_hsg <- filter(df_cr_m, Method == "HSG")$Median)
(filter(df_cr_m, Method == "HSG")$Max)

# At which q does the SMLR median accuracy surpass that of the HSG ('accuracy vs q'-plot)?
# We used smoothed line in the plot, and we can compute this line by
d_smlr_med <- acc |> filter(Line == "Median")
loess.data <- stats::loess(Accuracy ~ q, data = d_smlr_med)
loess.predict <- predict(loess.data, se = T)
loess.df <- data.frame("Smoothed_accuracy" = loess.predict$fit, se = loess.predict$se.fit, "Data_accuracy" = d_smlr_med$Accuracy, q = d_smlr_med$q)
(a_med <- loess.df |> filter(Smoothed_accuracy >= a_med_hsg) |> filter(q==min(q)))
# (a_max <- acc |> filter(Line == "Median", Accuracy >= filter(df_acc_m, Method == "HSG")$Max) |> filter(q==min(q)))

# What call rate does this q-value translate into when taken to the 'call rate vs q'-plot?
(c_med <- cr |> filter(Line == "Median", q == plyr::round_any(a_med$q, 0.005)))
# (c_max <- cr |> filter(Line == "Median", q == a_max$q))

# This is an absolute increase of
c_med$Call_rate - c_med_hsg
plyr::round_any(c_med$Call_rate - c_med_hsg, 0.1)
# And a relative increase of
(c_med$Call_rate - c_med_hsg) / c_med_hsg

# What is the largest q-value for which the SMLR has the same call rate as the HSG or higher?
(q_cr <- cr |> filter(Line == "Median", Call_rate >= c_med_hsg) |> filter(q == max(q)))

# What accuracy does this q-value translate into when taken back to the 'accuracy vs q'-plot?
(q_ac <- loess.df |> filter(q == q_cr$q))

# This is an absolute increase of
plyr::round_any(q_ac$Smoothed_accuracy - a_med_hsg, 0.001)
# And a relative increase of
(q_ac$Smoothed_accuracy - a_med_hsg) / a_med_hsg

# (a_med$Smoothed_accuracy - a_med_hsg)
# (a_med$Smoothed_accuracy - a_med_hsg) / a_med_hsg
# (c_max$Call_rate - c_max_hsg) / c_max_hsg

# (c_max2 <- cr |> filter(Line == "Median", Call_rate >= c_max_hsg) |> filter(q==max(q)))
# (a_med2 <- acc |> filter(Line == "Median", q == c_max2$q))
# (a_max_hsg <- filter(df_acc_m, Method == "HSG")$Max)
# (a_med2$Accuracy - a_max_hsg) / a_max_hsg
```



# Accuracy vs call rate (all models)
```{r}
dd_I <- readRDS(file.path("..", "data", "01_obj_crossval_I_1K_9individuals_low_high_all.rds"))
dd_I_NI <- readRDS(file.path("..", "data", "01_obj_crossval_I_NI_1K_9individuals_low_high_all.rds"))
dd_log <- readRDS(file.path("..", "data", "01_obj_crossval_log_1K_9individuals_low_high_all.rds"))
dd_log_NI <- readRDS(file.path("..", "data", "01_obj_crossval_log_NI_1K_9individuals_low_high_all.rds"))
dd_sqrt <- readRDS(file.path("..", "data", "01_obj_crossval_sqrt_1K_9individuals_low_high_all.rds"))
dd_sqrt_NI <- readRDS(file.path("..", "data", "01_obj_crossval_sqrt_NI_1K_9individuals_low_high_all.rds"))

dd_all <- list("I" = dd_I, "I_NI" = dd_I_NI,
               "log" = dd_log, "log_NI" = dd_log_NI,
               "sqrt" = dd_sqrt, "sqrt_NI" = dd_sqrt_NI)
# Restructure list of lists (to make the subsequent script work more natural):
dd_all <- lapply(names(dd_all[[1]]), function(name) {
  lapply(dd_all, function(x) x[[name]])
})
names(dd_all) <- c("low", "high", "very_high")
```

```{r}
dd_models <- lapply(dd_all, function(d) {
  lapply(d, function(m) {
  dd_r <- lapply(m, function(r) {
    dd_pmax <- data.frame("p_max" = c(0, r$p_WC, r$p_NN),
                          "WC_l" = c(F, rep(T, length(r$p_WC)), rep(F, length(r$p_NN)))) |> 
      arrange(p_max) |> 
      mutate(WC = length(r$p_WC) - cumsum(WC_l), NC = 0:r$N_test)
    last_row <- dd_pmax[nrow(dd_pmax),] |> # to prevent division by zero
      mutate(Accuracy = 100,
             Call_rate = 0)
    dd_pmax <- dd_pmax[-nrow(dd_pmax),] |> 
      mutate(Accuracy = 100*(r$N_test - NC - WC) / (r$N_test - NC),
             Call_rate = 100*(r$N_test - NC) / r$N_test) |> 
      bind_rows(last_row) |> 
      mutate(CR_bin = round(Call_rate, digits = 1))
    
    return(dd_pmax)
  }) |> bind_rows()
  dd_r <- dd_r |> group_by(CR_bin) |> 
    summarise(Mean_AC = mean(Accuracy),
              Median_AC = median(Accuracy),
              Mean_CR = mean(Call_rate),
              Median_CR = median(Call_rate),
              Bin_count = n())
  
  return(dd_r)
}) |> bind_rows(.id = "Model")
})
```

```{r}
HSG_WC_NC <- lapply(dd_all, function(d) {
  # The HSG has the same results for all models, so we can access them from e.g. the square-root model
  dd_hsg <- lapply(d$sqrt, function(r) {
    data.frame("Accuracy" = 100*(r$N_test - r$HSG_NN - r$HSG_WC) / (r$N_test - r$HSG_NN),
               "Call_rate" = 100*(r$N_test - r$HSG_NN) / r$N_test,
               "Model" = "HSG")
    }) |> bind_rows() |> 
    group_by(Model) |> 
    summarise(Accuracy = median(Accuracy),
              Call_rate = median(Call_rate))
  })
```

```{r}
(p_low <- ggplot(dd_models$low, aes(x = CR_bin, y = Median_AC, colour = Model)) +
  geom_line(show.legend = F) +
  geom_point(aes(x = Call_rate, y = Accuracy, colour = Model), data = HSG_WC_NC$low, show.legend = F) +
  coord_cartesian(xlim = c(100, 40), ylim = c(99, 100)) +
  scale_x_reverse() +
  xlab("Call rate [%]") + ylab("Median accuracy [%]") + 
  ggtitle("Fit: 25pg, 50pg\nTest: 31.25pg, 50pg") +
  theme_bw() +
  theme(plot.title = element_text(size = plot_title_size),
        axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_ticks_size),
        legend.position = "none"))

(p_high <- ggplot(dd_models$high, aes(x = CR_bin, y = Median_AC, colour = Model)) + 
  geom_line() + 
  geom_point(aes(x = Call_rate, y = Accuracy, colour = Model), data = HSG_WC_NC$high) +
  guides(colour = guide_legend(position = "inside")) +
  coord_cartesian(xlim = c(100, 95), ylim = c(99.95, 100)) +
  scale_x_reverse() +
  xlab("Call rate [%]") + ylab("Median accuracy [%]") + 
  ggtitle("\n62.5pg, 125pg, 250pg") +
  theme_bw() +
  theme(plot.title = element_text(size = plot_title_size),
        axis.title = element_text(size = axis_title_size),
        axis.text = element_text(size = axis_ticks_size),
        legend.title = element_text(size = legend_title_size),
        legend.text = element_text(size = legend_text_size),
        legend.position.inside = c(0.837,0.48),
        legend.key.spacing = unit(3, 'mm'),
        legend.key.height = unit(5, 'mm'),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black")))

# p_very_high <- ggplot(dd_models$very_high, aes(x = CR_bin, y = Median_AC, colour = Model)) + 
#   geom_line(show.legend = F) + 
#   geom_point(aes(x = Call_rate, y = Accuracy, colour = Model), data = HSG_WC_NC$very_high, show.legend = F) +
#   coord_cartesian(xlim = c(100, 99.9), ylim = c(99.95, 100)) +
#   scale_x_reverse(breaks = c(100, 99.95, 99.9), labels = c("100", "99.95", "99.9")) +
#   xlab("Call rate [%]") + ylab("Median accuracy [%]") + 
#   ggtitle("\n500pg, 1000pg") +
#   theme_bw() +
#   theme(plot.title = element_text(size = plot_title_size),
#         axis.title = element_text(size = axis_title_size),
#         axis.text = element_text(size = axis_ticks_size),
#         legend.title = element_text(size = legend_title_size),
#         legend.text = element_text(size = legend_text_size))
```

```{r}
# (p <- p_low + p_high + p_very_high +
#    plot_layout(ncol = 3, nrow = 1, byrow = T, guides = "collect", axis_titles = "collect") &
#    scale_colour_manual(values=cbp1[c(7, 1:6)],
#                        breaks = c("HSG", "I", "I_NI", "log", "log_NI", "sqrt", "sqrt_NI"),
#                        labels = unname(c("HSG",
#                                          TeX("$I$"),
#                                          TeX("$I, \\beta_0 = 0$"),
#                                          TeX("$log$"),
#                                          TeX("$log, \\beta_0 = 0$"),
#                                          TeX("$sqrt$"),
#                                          TeX("$sqrt, \\beta_0 = 0$")))))

p <- p_low + p_high +
   plot_layout(ncol = 2, nrow = 1, byrow = T, axis_titles = "collect") &
   scale_colour_manual(values=cbp1[c(7, 1:6)],
                       breaks = c("HSG", "I", "I_NI", "log", "log_NI", "sqrt", "sqrt_NI"),
                       labels = unname(c("HSG",
                                         TeX("$I$"),
                                         TeX("$I, \\beta_0 = 0$"),
                                         TeX("$log$"),
                                         TeX("$log, \\beta_0 = 0$"),
                                         TeX("$sqrt$"),
                                         TeX("$sqrt, \\beta_0 = 0$")))) &
  theme(plot.margin = margin(t=0, r=5.5, b=0, l=5.5, unit = 'pt'))
p[[1]]$theme$plot.margin[4] <- 0*p[[1]]$theme$plot.margin[4]
p[[2]]$theme$plot.margin[2] <- 0*p[[2]]$theme$plot.margin[2]
(p <- p + plot_annotation(theme = theme(plot.margin = margin(t=0, r=0, b=0, l=0, unit = 'pt'))))

ggsave(filename = file.path("..", "article-FSIGEN", "FIG_3-two-column.bmp"),
       plot = p, device = "bmp", width = 7480, height = 2480, units = "px", dpi = 1000)
```

*NOTE*: below a call rate of $50\%$ in the left plot (corresponding to q-values above 0.999), the yellow line (model with intercept and square root transformation) jumps a bit up and down.
At first glance, this may seem odd, but it's not!
Remember: the first axis was divided into call rate bins (with a width of one per cent) and within each bin, the median accuracy was computed.
There is nothing odd in that the median accuracy goes a little up and down for a few consecutive bins.
Had we chosen a more coarse first axis (larger call rate bin size), the jumping would likely be smoothed out.
However, this would have resulted in a less accurate depiction at the call rates above $90\%$.
After all, the higher call rates are the most interesting, so we want a fine call rate scale in that range to be able to distinguish the near vertical lines in the left side of the plot.
Therefore, we accept the side effect of a little jumping for this single line in the less interesting right side of the plot --- we could easily argue that it would be sufficient to only let the first axis extend down to a call rate of $50\%$ (or even higher), which would remove the part with the jumping line.

The creation of the plot of the median accuracy vs. call rate with the HSG's performance represented as a red dot was guided by the fact that the HSG has just one call rate and accuracy per cross validation iteration, whereas each SMLR model has a variable call rate and accuracy that ultimately depends on the probability threshold, q.
Thus, each SMLR model's median accuracy within each bin of call rates will be based on result across all cross-validation iterations.
However, if we tried to make call rate bins for the HSG, these would not be able to contain accuracies from all cross-validation iterations, unless all accuracies were observed within the same bin, in which case the result would be a point very similar to the red dot from above.
The intuition behind that visualization approach is, that it is best to base the displayed points and lines on all cross-validation iterations.

But let's also try to see what happens, if we make call rate bins for the HSG:
```{r}
HSG_WC_NC_bins <- lapply(dd_all, function(d) {
  m <- d$sqrt
  dd_r <- lapply(m, function(r) {
    dd_hsg <- data.frame("Accuracy" = 100*(r$N_test - r$HSG_NN - r$HSG_WC) / (r$N_test - r$HSG_NN),
                         "Call_rate" = 100*(r$N_test - r$HSG_NN) / r$N_test) |> 
      mutate(CR_bin = round(Call_rate, digits = 1))
    return(dd_hsg)
  }) |> bind_rows()
  dd_r <- dd_r |> group_by(CR_bin) |> 
    summarise(Mean_AC = mean(Accuracy),
              Median_AC = median(Accuracy),
              Mean_CR = mean(Call_rate),
              Median_CR = median(Call_rate),
              Bin_count = n())
  
  return(cbind("Model" = "HSG", dd_r))
})
```

```{r}
p_low <- ggplot(dd_models$low, aes(x = CR_bin, y = Median_AC, colour = Model)) +
  geom_line() +
  geom_point(aes(x = CR_bin, y = Median_AC, colour = Model), data = HSG_WC_NC_bins$low) +
  coord_cartesian(xlim = c(100, 40), ylim = c(99, 100)) +
  scale_x_reverse() +
  xlab("Call rate [%]") + ylab("Median accuracy [%]") + 
  ggtitle("Fit: 25pg, 50pg\nTest: 31.25pg, 50pg") +
  # theme_bw() +
  theme(axis.title.y=element_text(size = axis_title_size), plot.title = element_text(size = plot_title_size))

p_high <- ggplot(dd_models$high, aes(x = CR_bin, y = Median_AC, colour = Model)) + 
  geom_line(show.legend = F) + 
  geom_point(aes(x = CR_bin, y = Median_AC, colour = Model), data = HSG_WC_NC_bins$high, show.legend = F) +
  coord_cartesian(xlim = c(100, 95), ylim = c(99.95, 100)) +
  scale_x_reverse() +
  xlab("Call rate [%]") + ylab("Median accuracy [%]") + 
  ggtitle("\n62.5pg, 125pg, 250pg") +
  # theme_bw() +
  theme(axis.title.y=element_blank(),
        plot.title = element_text(size = plot_title_size))
```

```{r}
(p <- p_low + p_high +
   plot_layout(ncol = 2, nrow = 1, byrow = T, guides = "collect", axis_titles = "collect") &
   scale_colour_manual(values=cbp1[c(7, 1:6)],
                       breaks = c("HSG", "I", "I_NI", "log", "log_NI", "sqrt", "sqrt_NI"),
                       labels = unname(c("HSG",
                                         TeX("$I$"),
                                         TeX("$I, \\beta_0 = 0$"),
                                         TeX("$log$"),
                                         TeX("$log, \\beta_0 = 0$"),
                                         TeX("$sqrt$"),
                                         TeX("$sqrt, \\beta_0 = 0$")))))

ggsave(filename = file.path("..", "article-FSIGEN", "accuracy_vs_call_rate_bins.bmp"), plot = p,
       device = "bmp", width = 30, height = 10.5, units = "cm")
```

The result of this visualization test is:
The HSG has a rather constant accuracy and call rate across all cross-validation iterations, so we don't really lose any information by displaying a single point.
For the higher DNA amounts (62.5pg, 125pg, and 250pg) we actually only got one call rate bin, so here it was only possible to display a single point (when rounding call rates to one decimal).
For the lower DNA amounts, we got nine call rate bins ranging from 93.2 to 94.0, and they all had very close median accuracies:
```{r}
table(round(HSG_WC_NC_bins$low$Median_AC, 3))
```
where the highest median accuracy was for a bin that only contained three observations/cross-validation iterations.
```{r}
HSG_WC_NC_bins$low[,c("Bin_count", "Median_AC")]
```
This was exactly what I wanted to avoid, when I insisted on "that it is best to base the displayed points and lines on all cross-validation iterations".

The minimum observed points in SMLR models' call rate bins for the ranges of DNA quantities are
```{r}
sapply(dd_models, function(d){min(d$Bin_count)})
```
so their median accuracies are definitely well estimated.


# Summary statistics to be reported in article
```{r}
# Low dilutions (31.25pg and 50pg)
# HSG median accuracy and call rate
(hsg_med_ac_cr_low <- round(HSG_WC_NC$low[,c("Accuracy", "Call_rate")], digits = 3))

# When does SMLR surpass HSG's median accuracy and what is its accuracy at call rate 95%
SMLR_med_ac_cr_low <- dd_models$low |> filter(Model == "sqrt") |> select(CR_bin, Median_AC, Bin_count)
SMLR_med_ac_cr_low |> filter(Median_AC >= hsg_med_ac_cr_low$Accuracy) |> 
  filter(CR_bin == max(CR_bin))
SMLR_med_ac_cr_low |> filter(CR_bin == 95)


# High dilutions (62.5pg, 125pg, and 250pg)
# HSG median accuracy and call rate
(hsg_med_ac_cr_high <- round(HSG_WC_NC$high[,c("Accuracy", "Call_rate")], digits = 3))

# When does SMLR surpass HSG's median accuracy and what is its accuracy at call rate 95%
SMLR_med_ac_cr_high <- dd_models$high |> filter(Model == "sqrt") |> select(CR_bin, Median_AC, Bin_count)
SMLR_med_ac_cr_high |> filter(Median_AC >= hsg_med_ac_cr_high$Accuracy) |> 
  filter(CR_bin == max(CR_bin))
SMLR_med_ac_cr_high |> filter(Median_AC == 100) |> 
  filter(CR_bin == max(CR_bin))


# Very high dilutions (500pg, 1000pg)
# HSG median accuracy and call rate
(hsg_med_ac_cr_very_high <- round(HSG_WC_NC$very_high[,c("Accuracy", "Call_rate")], digits = 3))

# When does SMLR surpass HSG's median accuracy and what is its accuracy at call rate 95%
SMLR_med_ac_cr_very_high <- dd_models$very_high |> filter(Model == "sqrt") |> select(CR_bin, Median_AC, Bin_count)
SMLR_med_ac_cr_very_high |> filter(Median_AC >= hsg_med_ac_cr_very_high$Accuracy) |> 
  filter(CR_bin == max(CR_bin))
SMLR_med_ac_cr_very_high |> filter(CR_bin == 95)
```



# HIGH
# Accuracy and call rate vs. q for the cv winner fit testing on high dilutions
```{r}
# dd_val_dils <- readRDS(file.path("..", "data", "01_obj_crossval_sqrt_NI_1K_9individuals_low_high_all.rds"))
dd_val_dils <- readRDS(file.path("..", "data", "01_obj_crossval_sqrt_1K_9individuals_low_high_all.rds"))
dd_val_dils <- dd_val_dils[2]
names(dd_val_dils) <- "D"
```

```{r}
# df_acc <- lapply(dd_val_dils$D, function(x) {
#   q <- floor_dec(x$p_WC, 4)
#   data.frame("q" = c(q, max(q) + 0.0001),
#              "Accuracy" = c(((x$N_test - rev(1:length(q))) / x$N_test)*100, 100))
# }) |> bind_rows(.id = "Line")
df_acc <- lapply(dd_val_dils$D, function(x) {
  q <- c(x$p_WC, min(x$p_NN[x$p_NN > max(x$p_WC)]))
  WC <- sapply(q, function(y) {sum(x$p_WC >= y)})
  p_all <- c(x$p_WC, x$p_NN)
  NC <- sapply(q, function(y) {sum(p_all < y)})
  
  q <- floor_dec(q, 4)
  a <- floor_dec(((x$N_test - WC - NC) / (x$N_test - NC))*100, 2)
  data.frame("q" = q,
             "Accuracy" = a)
}) |> bind_rows(.id = "Line") ##|> 
  # mutate(Line = as.factor((80+(as.numeric(Line)-1)*(100-80)/(1000-1))/100))


acc <- unique(df_acc$Accuracy)[order(unique(df_acc$Accuracy))]
qq_acc <- matrix(NA, nrow = length(acc), ncol = 3)
qq_acc[1,] <- quantile(filter(df_acc, Accuracy == min(df_acc$Accuracy))$q, probs = c(q_l, 0.5, q_u))
qq_acc[length(acc),] <- quantile(filter(df_acc, Accuracy == max(df_acc$Accuracy))$q, probs = c(q_l, 0.5, q_u))
for (i in 2:(length(acc)-1)) {
  qq_acc[i,] <- quantile(filter(df_acc, Accuracy == acc[i])$q, probs = c(q_l, 0.5, q_u))
}
colnames(qq_acc) <- names_col
qq_acc <- as.data.frame(qq_acc)
qq_acc$Accuracy <- acc
qq_acc <- qq_acc |> pivot_longer(cols = !Accuracy, names_to = "Line", values_to = "q") |> 
  mutate(Method = "SMLR") |> 
  select(Method, q, Line, Accuracy)

df_acc_m <- lapply(dd_val_dils$D, function(x) {
  data.frame("Accuracy_HSG" = ((x$N_test - x$HSG_WC - x$HSG_NN) / (x$N_test - x$HSG_NN))*100,
             "Accuracy_EQC" = ((x$N_test - x$EQC_WC - x$EQC_NN) / (x$N_test - x$EQC_NN))*100)
}) |> bind_rows(.id = "Line")
df_acc_m <- data.frame("Accuracy" = c(min(df_acc_m$Accuracy_HSG),
                                      quantile(df_acc_m$Accuracy_HSG, probs = c(q_l, 0.5, q_u)),
                                      max(df_acc_m$Accuracy_HSG),
                                      min(df_acc_m$Accuracy_EQC),
                                      quantile(df_acc_m$Accuracy_EQC, probs = c(q_l, 0.5, q_u)),
                                      max(df_acc_m$Accuracy_EQC)),
                       "Method" = rep(c("HSG", "EQC"), each=5),
                       "Line" = rep(c("Min", names_col, "Max"), 2),
                       "x1" = 0.5,
                       "x2" = 0.8) |> 
  pivot_wider(values_from = "Accuracy", names_from = "Line")
df_acc_stats <- df_acc_m |> select(Method, Q25, Median, Q75, x1, x2) |> 
  # pivot_longer(cols = c(x1, x2), values_to = "q", names_to = NULL) |> 
  pivot_longer(cols = c(Q25, Median, Q75), values_to = "Accuracy", names_to = "Line")


(p_acc <- ggplot() + 
  # geom_line(aes(x=q, y=Accuracy, alpha=Line), colour="grey", show.legend = F, data = df_acc) +
  geom_line(aes(x=q, y=Accuracy, colour=Line), show.legend = F, data = df_acc) +
  scale_colour_manual(values = rep("grey",1000)) +
  new_scale_color() +
  geom_rect(aes(xmin=x1, xmax=x2, ymin=Min, ymax=Max, fill=Method), data=df_acc_m, alpha=0.3, show.legend = F) +
  scale_fill_manual(breaks=c("HSG", "EQC"),
                    values= c("#F8766D", "#619CFF"),
                    labels=c("HSG", "EQC")) +
  
  # geom_rect(aes(xmin=x1, xmax=x2, ymin=Q25, ymax=Q75, fill=Method), data=df_acc_m, alpha=0.6) +
  geom_line(aes(x=q, y=Accuracy, lty=Line, colour=Method), data = qq_acc) +
  geom_segment(aes(x=x1, y=Accuracy, xend=x2, yend=Accuracy, colour=Method, lty=Line), data=df_acc_stats) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "black"),
                      breaks = c("HSG", "EQC", "SMLR")) +
  scale_linetype_manual(values = c("dotted", "dashed", "dotted"),
                        breaks = c("Q75", "Median", "Q25")) +
  theme_bw() +
  ggtitle("Fit: 50pg, 25pg\nTest: 250pg, 125pg, 62.5pg") +
  theme(plot.title = element_text(size = 7),
        legend.position = "none"))

# ggsave(filename = file.path("..", "article-FSIGEN", "accuracy_vs_q_high.pdf"), plot = p_acc,
#        device = "pdf", width = 18, height = 11, units = "cm")
```


```{r}
df_cr <- lapply(dd_val_dils$D, function(x) {
  q <- c(x$p_WC, x$p_NN)
  q <- floor_dec(q[order(q)], 4)
  data.frame("q" = q,
             "Call_rate" = ((x$N_test - 0:(length(q) - 1)) / x$N_test)*100)
}) |> bind_rows(.id = "Line") ##|> 
  # mutate(Line = as.factor((80+(as.numeric(Line)-1)*(100-80)/(1000-1))/100))


qq <- seq(0.5, 0.95, 0.01)
qq2 <- seq(0.96, 1, 0.001)
cr <- matrix(NA, nrow = length(c(qq, qq2)), ncol = 3)
cr[1,] <- quantile(filter(df_cr, q < 0.505)$Call_rate, probs = c(q_l, 0.5, q_u))
cr[nrow(cr),] <- quantile(filter(df_cr, q >= 0.9995)$Call_rate, probs = c(q_l, 0.5, q_u))
for (i in 2:(length(qq))) {
  cr[i,] <- quantile(filter(df_cr, q >= qq[i]-0.005, q < qq[i]+0.005)$Call_rate, probs = c(q_l, 0.5, q_u))
}
for (i in 1:(length(qq2)-1)) {
  cr[length(qq)+i,] <- quantile(filter(df_cr, q >= qq2[i]-0.0005, q < qq2[i]+0.0005)$Call_rate, probs = c(q_l, 0.5, q_u))
}
colnames(cr) <- names_col
cr <- as.data.frame(cr)
cr$q <- c(qq, qq2)
cr <- cr |> pivot_longer(cols = !q, names_to = "Line", values_to = "Call_rate") |> 
  mutate(Method = "SMLR") |> 
  select(Method, q, Line, Call_rate)
cr <- cr |> group_by(Line) |> arrange(desc(Call_rate), .by_group = T) |> mutate(q = c(qq, qq2))


df_cr_m <- lapply(dd_val_dils$D, function(x) {
  data.frame("Call_rate_HSG" = ((x$N_test - x$HSG_NN) / x$N_test)*100,
             "Call_rate_EQC" = ((x$N_test - x$EQC_NN) / x$N_test)*100)
}) |> bind_rows(.id = "Line")
df_cr_m <- data.frame("Call_rate" = c(min(df_cr_m$Call_rate_HSG),
                                      quantile(df_cr_m$Call_rate_HSG, probs = c(q_l, 0.5, q_u)),
                                      max(df_cr_m$Call_rate_HSG),
                                      min(df_cr_m$Call_rate_EQC),
                                      quantile(df_cr_m$Call_rate_EQC, probs = c(q_l, 0.5, q_u)),
                                      max(df_cr_m$Call_rate_EQC)),
                      "Method" = rep(c("HSG", "EQC"), each=5),
                      "Line" = rep(c("Min", names_col, "Max"), 2),
                       "x1" = 0.5,
                       "x2" = 1) |> 
  pivot_wider(values_from = "Call_rate", names_from = "Line")
df_cr_stats <- df_cr_m |> select(Method, Q25, Median, Q75, x1, x2) |> 
  # pivot_longer(cols = c(x1, x2), values_to = "q", names_to = NULL) |> 
  pivot_longer(cols = c(Q25, Median, Q75), values_to = "Call_rate", names_to = "Line")



(p_cr <- ggplot() + 
  # geom_line(aes(x=q, y=Call_rate, alpha=Line), colour="grey", show.legend = F, data = df_cr) +
  geom_line(aes(x=q, y=Call_rate, colour=Line), show.legend = F, data = df_cr) +
  scale_colour_manual(values = rep("grey",1000)) +
  new_scale_color() +
  geom_rect(aes(xmin=x1, xmax=x2, ymin=Min, ymax=Max, fill=Method), data=df_cr_m, alpha=0.3, show.legend = F) +
  scale_fill_manual(breaks=c("HSG", "EQC"),
                    values= c("#F8766D", "#619CFF"),
                    labels=c("HSG", "EQC")) +
  # geom_rect(aes(xmin=x1, xmax=x2, ymin=Q25, ymax=Q75, fill=Method), data=df_acc_m, alpha=0.6) +
  geom_line(aes(x=q, y=Call_rate, lty=Line, colour=Method), data = cr) +
  geom_segment(aes(x=x1, y=Call_rate, xend=x2, yend=Call_rate, colour=Method, lty=Line), data=df_cr_stats) +
  scale_colour_manual(values = c("#F8766D", "#619CFF", "black"),
                      breaks = c("HSG", "EQC", "SMLR")) +
  scale_linetype_manual(values = c("dotted", "dashed", "dotted"),
                        breaks = c("Q75", "Median", "Q25")) +
  theme_bw() +
  # ggtitle("Fit: 50pg, 25pg\nTest: 250pg, 125pg, 62.5pg") +
  theme(plot.title = element_text(size = 10)) +
    ylab("Call rate"))

# ggsave(filename = file.path("..", "article-FSIGEN", "call-rate_vs_q_high.bmp"), plot = p_cr,
#        device = "bmp", width = 18, height = 11, units = "cm")
```

```{r}
(p <- p_acc + p_cr + plot_layout(ncol = 2, nrow = 1, byrow = T, guides = "collect"))
ggsave(filename = file.path("..", "article-FSIGEN", "accr_vs_q_high.bmp"), plot = p,
       device = "bmp", width = 30, height = 11, units = "cm")
```


### Summary statistics for high
```{r}
filter(df_acc_stats, Line == "Median", Method == "HSG")$Accuracy
(q100 <- qq_acc |> filter(Line == "Median", Accuracy == 100) |> filter(q==min(q)))
(cr |> filter(Line == "Median", q >= q100$q) |> filter(q==min(q)))
# filter(df_acc_m, Method == "HSG")$Max
# 
(a_med <- qq_acc |> filter(Line == "Median", Accuracy >= filter(df_acc_stats, Line == "Median", Method == "HSG")$Accuracy) |> filter(q==min(q)))
# (a_max <- qq_acc |> filter(Line == "Median", Accuracy >= filter(df_acc_m, Method == "HSG")$Max) |> filter(q==min(q)))
(c_med <- cr |> filter(Line == "Median", q < a_med$q) |> filter(q == max(q)))
# (c_max <- cr |> filter(Line == "Median", q == a_max$q))
(c_med_hsg <- filter(df_cr_m, Method == "HSG")$Median)
# (c_max_hsg <- filter(df_cr_m, Method == "HSG")$Max)
(c_med$Call_rate - c_med_hsg) / c_med_hsg
# (c_max$Call_rate - c_max_hsg) / c_max_hsg

# (c_max2 <- cr |> filter(Line == "Median", Call_rate >= c_max_hsg) |> filter(q==max(q)))
# (a_med2 <- acc |> filter(Line == "Median", q == c_max2$q))
# (a_max_hsg <- filter(df_acc_m, Method == "HSG")$Max)
# (a_med2$Accuracy - a_max_hsg) / a_max_hsg
# 
(a_med <- qq_acc |> filter(Line == "Median", Accuracy >= filter(df_acc_stats, Line == "Median", Method == "EQC")$Accuracy) |> filter(q==min(q)))
# (a_max <- qq_acc |> filter(Line == "Median", Accuracy >= filter(df_acc_m, Method == "EQC")$Max) |> filter(q==min(q)))
(c_med <- cr |> filter(Line == "Median", q < a_med$q) |> filter(q == max(q)))
# (c_max <- cr |> filter(Line == "Median", q == a_max$q))
(c_med_eqc <- filter(df_cr_m, Method == "EQC")$Median)
# (c_max_eqc <- filter(df_cr_m, Method == "EQC")$Max)
(c_med$Call_rate - c_med_eqc) / c_med_eqc
# (c_max$Call_rate - c_max_eqc) / c_max_eqc

# (c_max2 <- cr |> filter(Line == "Median", Call_rate >= c_max_eqc) |> filter(q==max(q)))
# (a_med2 <- qq_acc |> filter(Line == "Median", q == floor_dec(c_max2$q, 2)))
# (a_max_eqc <- filter(df_acc_m, Method == "EQC")$Max)
# (a_med2$Accuracy - a_max_eqc) / a_max_eqc

# At which q does the SMLR have the same call rate as the HSG
(q_hsg <- cr |> filter(Line == "Median", Call_rate >= c_med_hsg) |> filter(q == max(q)))
df_acc |> filter(Accuracy == 100, q >= q_hsg$q) |> nrow()
```